---
title: "Tooth Growth Data (Statistical Inference Course Project Part 2)"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---
by José Carrasquero

##1. Load the ToothGrowth data and perform some basic exploratory data analyses 
The following packages are used for this analysis:

```{r, warning=FALSE, message=FALSE}
library(graphics)
library(datasets)
```

```{r, warning=FALSE, cache = TRUE}
require(graphics)
coplot(len ~ dose | supp, data = ToothGrowth, panel = panel.smooth,
       xlab = "ToothGrowth data: length vs dose, given type of supplement")
##Box len ~ supp
boxplot(len ~ supp, ToothGrowth)

##Box len ~ dose
boxplot(len ~ dose, ToothGrowth)
```


##2. Provide a basic summary of the data.
```{r, warning=FALSE, cache = TRUE}
print(ToothGrowth)

summary(ToothGrowth)

#Different doses
unique(ToothGrowth$dose)
```

##3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)

```{r, warning=FALSE, cache = TRUE}
t.test(len ~ supp, ToothGrowth, paired = FALSE, var.equal = FALSE, conf.level = 0.95)

t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == 0.5, ], paired = FALSE, var.equal = FALSE, conf.level = 0.95)

t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == 1, ],paired = FALSE, var.equal = FALSE, conf.level = 0.95)

t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == 2, ],paired = FALSE, var.equal = FALSE, conf.level = 0.95)

t.test(len ~ dose, ToothGrowth[ToothGrowth$dose != 2, ], paired = FALSE, var.equal = FALSE, conf.level = 0.95)

t.test(len ~ dose, ToothGrowth[ToothGrowth$dose != 1, ], paired = FALSE, var.equal = FALSE, conf.level = 0.95)

t.test(len ~ dose, ToothGrowth[ToothGrowth$dose != 0.5, ], paired = FALSE, var.equal = FALSE, conf.level = 0.95)

```

##4. State your conclusions and the assumptions needed for your conclusions. 

Assumptions:

1. The Data is not Paired. It is a Two Sample T Test.

2. From the box plots we can assume the variences are not equal.

Conclusion:
1. Suplement type has no effect on tooth growth. (this is seen in the first 3 T test)
2. The dose level does have an effect on tooth growth. (this can be seen in the last 3 t test)